Sparse spectral and $p$-finite element methods on disk-slices and trapeziums I: zero boundary conditions

### 2019-04-08 papers/halfdisk/halfdisk.tex
New figures for paper

\begin{figure}
	\includegraphics[scale=0.3]{solution}
	\centering
	\caption{The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
        	\label{fig:solution-poisson}
	\centering
\end{figure}

@@

We can tackle the problem as follows. Fix \(a, b = 1,1\). Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above. We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matries. We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$. In matrix-vector notation, our system hence becomes:
\begin{align}
    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
\end{align}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in FIGURE with $v(x,y) = xy^2$ as an example. In FIGURE we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.

@@

Find \(u(x,y)\) given a function \(f(x,y)\) such that:
\begin{align}
	\begin{cases}
    		\Delta^2 u(x,y) = 0 \quad \text{in } \Omega \\
		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
	\end{cases}.
	\label{eqn:biharmonic}
\end{align}
where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$. We can construct a matrix operator for $\Delta^2$ that will take coefficients in the space
$$
\bigW^{(2,2)}
$$
to coefficients in the space
$$
\bigP^{(2,2)}.
$$
Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$. The matrix operator for the Biharmonic operator is then given by
\begin{align}
	\biharmonic = \laplaceoo \: \laplacewtt.
\end{align}
In FIGURE we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.

>>>>

\begin{figure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-poisson}
	\caption{}
	\label{fig:solution-poisson}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-helmholtz}
	\caption{}
	\label{fig:solution-helmholtz}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\centering
	\includegraphics[scale=0.3]{solution-biharmonic}
	\caption{}
	\label{fig:solution-biharmonic}
	\end{subfigure}
	\centering
    	\caption{(a): The computed solution to $\Delta u = f$ with zero boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$. (b): The computed solution to $\Delta u + k^2 \: v \: u = f$ with zero boundary conditions with $f(x,y) = xye^x$, $v(x,y) = xy^2$ and $k = 0.25$. (c): The computed solution to $\Delta^2 u = f$ with zero Dirichlet and Neumann boundary conditions with $f(x,y) = 1 + \text{erf}(5(1 - 10((x - 0.5)^2 + y^2)))$.}
        \label{fig:solution}
\end{figure}

@@

We can tackle the problem as follows. Fix \(a, b = 1,1\). Denote the coefficient vector for expansion of $u$ in the $\bigWii$ OP basis up to degree $N$ by $\mathbf{u}$, and the coefficient vector for expansion of $f$ in the $\bigPii$ OP basis up to degree $N$ by $\mathbf{f}$. Since $f$ is known, we can obtain $\mathbf{f}$ using the quadrature rule above. We can obtain the matrix operator for the variable-coefficient function $v(x,y)$ by using the Clenshaw algorithm with matrix inputs as the Jacobi matrices ${J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top$, yielding an operator matrix of the same dimension as the input Jacobi matries. We can denote the resulting operator acting on coefficients in the $\bigPoo$ space by $V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top)$. In matrix-vector notation, our system hence becomes:
\begin{align}
    (\laplacewii + k^2 T^{(0,0)\to(1,1)} \: V({J_x^{(0,0)}}^\top, {J_y^{(0,0)}}^\top) \: T_W^{(1,1)\to(0,0)}) \mathbf{u} = \mathbf{f}
\end{align}
which can be solved to find $\mathbf{u}$. We can see the sparsity and structure of this matrix system in Figure \ref{fig:sparsity} with $v(x,y) = xy^2$ as an example. In Figure \ref{fig:solution-helmholtz} we see the solution to the inhomogenous variable-coefficient Helmholtz equation with zero boundary conditions given in (\ref{eqn:helmholtz}) in the half-disk $\Omega$, with  $v(x,y) = xy^2$ and $f(x,y) = xy e^x$.

@@

Find \(u(x,y)\) given a function \(f(x,y)\) such that:
\begin{align}
	\begin{cases}
    		\Delta^2 u(x,y) = 0 \quad \text{in } \Omega \\
		u(x,y) = 0, \quad \frac{\partial u}{\partial n}(x,y) = 0 \quad \text{on } \partial \Omega
	\end{cases}.
	\label{eqn:biharmonic}
\end{align}
where $\Delta^2$ is the Biharmonic operator, noting the imposition of zero Dirichlet and Neumann boundary conditions on $u$. We can construct a matrix operator for $\Delta^2$ that will take coefficients in the space
$$
\bigW^{(2,2)}
$$
to coefficients in the space
$$
\bigP^{(2,2)}.
$$
Note that any function expanded in the $\bigW^{(2,2)}$ basis will satisfy both zero Dirichlet and zero Neumann boundary conditions on $\Omega$. The matrix operator for the Biharmonic operator is then given by
\begin{align}
	\biharmonic = \laplaceoo \: \laplacewtt.
\end{align}
In Figure \ref{fig:solution-biharmonic} we see the solution to the Biharmonic equation (\ref{eqn:biharmonic}) in the half-disk $\Omega$.
